---
title: "Sakshi_31A_Project_2_Time Series Analysis of Global Energy Consumption"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

**Introduction:**

This data contains the Monthly data of energy Production, Consumption,
Imports, Exports, Stock exchanges and production and consumption from
different resources like renewables, nuclear and fossil fuels from 1973
to 2022.

Data was collected from the website: U.S. Energy Information
Administration. All the information is in the Quadrillion British
Thermal Unit

**Objective:**

To predict the future global energy consumption via primary source of
energy based on past energy consumption trends.

# Forecast using ARIMA Method:

```{r}
library(flexdashboard)
library(tidyverse)
library(shiny)
library(DT)
library(forecast)
library(quantmod)
library(tseries)
library(tsibble)
library(feasts)
library(rugarch)
library(FinTS)
```

# Loanding and cleaning of data

```{r}
mydata <- read.csv("D:/IIFT Trimester 3/RTSM_Project_2/Data/World Energy Overview.csv")

x <- as.Date(mydata$Date, '%d-%m-%Y')
year <- as.numeric(format(x,'%Y'))
month <- as.numeric(format(x,'%m'))
day <- as.numeric(format(x,'%d'))

mydata2 <- cbind(mydata,year,month,day)
View(mydata2)
```

# Total Energy Consumption over years

```{r}
plot(mydata2$year, mydata2$Total.Primary.Energy.Consumption,type='l',xlab='Year',ylab = 'Consumption',
     lty=2,xlim = c(1980,2021), ylim=c(5,10))
```

# Decomposition of data for Total Energy Consumption

```{r}
x1 = ts(mydata2$Total.Primary.Energy.Consumption, start=c(1980, 5), freq= 12)

tsr = decompose(x1) # tsr : Trend | Seasonality | Randomness
plot(tsr)
```

# ADF Test: Total Consumption over years:

**Objective:** To check if the data is stationary or not

```{r}

summary(mydata2$Total.Primary.Energy.Consumption)
adf.test(mydata2$Total.Primary.Energy.Consumption) # Time series is stationary

```

**Analysis:**

Method: Augmented Dicky Fuller Test for stationarity Null Hypothesis:
Data is not stationary Alternate Hypothesis: Data is stationary

here, p-value: 0.01 \< 0.05 hence we reject the null hypothesis

**Conclusion:** Data is stationary

**Implication:** Since, data is stationary we can move ahead for time
series analysis.

# Ljung-Box Test: Consumption over Years

**Objective:** To check whether data is auto-correlated or not

```{r}
lb_test_wn = Box.test(mydata2$Total.Primary.Energy.Consumption)
lb_test_wn # Time-series is auto-correlated

```

**Analysis:**

Method: Ljung-Box test for autocorrelation Null Hypothesis: Time series
data is not auto-correlated Alternate Hypothesis: Time series data is
auto-correlated

here, p-value= 2.2e-16 \< 0.05 hence, we reject the null hypothesis

**Conclusion:** Times series data is auto-correlated

**Implication:** since data is auto-correlated, we can go for
time-series analysis of data instead of regression

# ACF and PACF plots for Total Energy Consumption

**Objective:** To identify and quantify the autocorrelation patterns in
a time series, assisting in the selection of appropriate models for time
series analysis and forecasting.

```{r}
acf(mydata2$Total.Primary.Energy.Consumption)
pacf(mydata2$Total.Primary.Energy.Consumption)
```

**Analysis:**

ACF:

The ACF appears to show a trend of increasing correlation as the lag
increases. This means that the total primary energy consumption in a
given year is positively correlated with the consumption in previous
years.

PACF:

Lag 1: The PACF value is close to 0, indicating that there is no
significant correlation between the total primary energy consumption in
a given year and the consumption in the previous year, after accounting
for the trend.

Lag 2 and beyond: The PACF values are close to 0, suggesting that there
is no significant correlation between the total primary energy
consumption in a given year and the consumption in years further in the
past, after accounting for the trend and the previous year's
consumption.

**Implication:**

PACF plot suggests that the total primary energy consumption in a given
year is not significantly related to the consumption in previous years,
after considering the trend and the influence of the previous year's
consumption. This is in contrast to the ACF plot, which showed a
positive correlation at all lags.

# ARIMA Model

**Objective:** To determine the appropriate orders of autoregressive
(AR) and moving average (MA) components and to find out how closely
related manual and auto-ARIMA models are.

```{r}
x1 = ts(mydata2$Total.Primary.Energy.Consumption, start=c(1980, 5), freq= 12)
newData <- diff(x1)
# ARIMA (1, 0, 0) or AR(1)
ar1 = arima(x1, order = c(1, 0, 0)); ar1

# ARIMA (2, 0, 0) or AR(2)
ar2 = arima(x1, order = c(2, 0, 0)); ar2

# ARIMA (0, 0 , 1) or MA(1)
ma1 = arima(x1, order = c(0, 0, 1)); ma1

# ARIMA (0, 0, 2) or MA(2)
ma2 = arima(x1, order = c(0, 0, 2)); ma2

# ARIMA (0, 0, 3) or MA(3)
ma3 = arima(x1, order = c(0, 0, 3)); ma3

# ARIMA (0, 0, 4) or MA(4)
ma4 = arima(x1, order = c(0, 0, 4)); ma4

# ARIMA (1, 0, 1) or ARMA(1, 1)
arma11 = arima(x1, order = c(1, 0, 1)); arma11

# ARIMA (1, 0, 2) or ARMA(1, 2)
arma12 = arima(x1, order = c(1, 0, 2)); arma12

# ARIMA (1, 0, 3) or ARMA(1, 3)
arma13 = arima(x1, order = c(1, 0, 3)); arma13

# Auto ARIMA
arma_pq_consumption_ds = auto.arima(x1); arma_pq_consumption_ds

```

**Analysis:**

The positive AR coefficient suggests a positive autocorrelation,
indicating a trend. The negative MA coefficients suggest a corrective
effect, dampening the impact of past forecast errors. The negative
seasonal AR and MA coefficients indicate a seasonally adjusted model.

**Implication:**

The ARIMA model with the provided coefficients and statistics seems to
capture autocorrelation, seasonality, and a linear trend. The model fit
and goodness-of-fit criteria (AIC, AICc, BIC) suggest that this model
provides a reasonable balance between explanatory power and complexity.

The lowest AIC by manual ARIMA method is for order (1,0,0) and (1,0,2)
The auto-ARIMA model also suggests the order (1,0,2) and (2,1,1) for
seasonality. Hence manual and auto ARIMA models give nealr same order.

# Forecasting with ARIMA model

**Objective:** To forecast the future global energy consumption based on
past consumption

```{r}
# Ljung-Box Test for Autocorrelation - Model Residuals
# ****************************************************
lb_test_arma_pq_jj_ds = Box.test(arma_pq_consumption_ds$residuals); lb_test_arma_pq_jj_ds
# It is not autocorrelated

# 3.1.2. Forecasting with ARIMA Models
# ************************************
jj_ds_fpq = forecast(arma_pq_consumption_ds, h = 100)
plot(jj_ds_fpq)

```

**Analysis:**

Method: Ljung-Box test for autocorrelation Null Hypothesis: Time series
data residual is not auto-correlated Alternate Hypothesis: Time series
data residual is auto-correlated

here, p-value= 0.9468 \> 0.05 hence, we do not reject the null
hypothesis

**Conclusion:**

Residual is not auto correlated.

**Implication:**

Model forecasted the data for next 100 time points

# ARCH-LM Test for residual

**Objective:** To find out whether there is heteroskedaskity is there in
residual or not

```{r}
ts_data <- ts(mydata$Total.Primary.Energy.Consumption, frequency = 12)

nse_ret_arch_test = ArchTest(arma_pq_consumption_ds$residuals, lags = 20) # H0: No ARCH Effects
nse_ret_arch_test
```

**Analysis:**

Method: ARCH-LM Test Null Hypothesis: There is no evidence of
conditional heteroskedasticity in the time series residual. Alternate
Hypothesis: There is evidence of heteroskedasticity

here, p-value = 2.2e-16 \< 0.05 Hence we reject the null hypothesis

**Conclusion:** There is heteroskedasticity in the residual of model

**Implication:**

The result implies that the conditional variance of the time series
residual is not constant over time; instead, it exhibits autoregressive
conditional heteroskedasticity.This may suggest that the volatility of
the residual of series is not constant, and there are periods of high
and low volatility.

Hence we go for GARCH model

# GARCH Model

```{r}
garch_model1 = ugarchspec(variance.model = list(model = 'sGARCH', garchOrder = c(1,1)), mean.model = list(armaOrder = c(1,2), include.mean = TRUE))
nse_ret_garch1 = ugarchfit(garch_model1, data = arma_pq_consumption_ds$residuals); nse_ret_garch1
```

**Analysis:**

Method: ARCH-LM Test Null Hypothesis: There is no evidence of
conditional heteroskedasticity in the time series residual. Alternate
Hypothesis: There is evidence of heteroskedasticity

p-value \> 0.05 for ARMA order of (1,2) for group of 30

**Conclusion:**

There is no heteroskedasticity in the residual of model for ARMA order
of (1,2) and group of 30

# GARCH Forecast

**Objective:** To forecast volatility using the fitted GARCH model for
the next 30 time points. Analysis: Used the 'ugarchforecast' function to
generate volatility forecasts for the next 30 time points.

```{r}
stk_ret_garch_forecast1 = ugarchforecast(nse_ret_garch1, n.ahead = 30); stk_ret_garch_forecast1

```

Results:

GARCH Model Forecast: - Model: sGARCH - Horizon: 30 - Roll Steps: 0 -
Out of Sample: 0

0-roll forecast [T0=Mar 2030]: - Forecasted Series: - T+1 to T+30:
Contains forecasted values of volatility (Sigma) for each time point.

Implication:

The forecasted values represent the predicted volatility for the next 30
time points based on the fitted GARCH model. These forecasts can be
useful for operations management and decision-making, providing insights
into the expected future volatility of the consumption time series.

# GARCH Forecast Plot

```{r}
plot(stk_ret_garch_forecast1)
```
